---
title: "Practical Machine Learning - Course Project"
author: "stern0phant"
date: "30 April 2016"
output: html_document
---

## 1 Introduction
This report was created for the peer assessment of the Practical Machine Learning course project on Coursera. The project's goal is to utilize provided data to build a model, which utilizes accelerometer data to evaluate in which way a physical exercise is performed.

## 2 Model Building
The following sub-sections describe the process of building the prediction model.

### 2.1 Preparations
```{r Load packages, message=FALSE}
## Load additional packages
library(caret)
library(dplyr)
library(randomForest)
```

```{r Load data, cache=TRUE}
## Load provided data
download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",
              destfile = "data.csv",
              method = "auto")

download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",
              destfile = "data_submit.csv",
              method = "auto")

data <- read.csv("data.csv", na.strings = c("NA", "", "#DIV/0!"))
data_submit <- read.csv("data_submit.csv")

## Create a training and a test sample
set.seed(741)
inTrain = createDataPartition(data$X, p = 0.80, list = F)
data_train <- data[inTrain, ]
data_test  <- data[-inTrain, ]
```

### 2.2 Predictor selection
```{r}
predictors <- names(data_train)

predictors <- predictors[!(predictors == "classe")] # Remove outcome variable
predictors <- predictors[!(predictors == "X")] # Remove count variable

# Remove predictors, which are frequently NA
for (i in 1:160) {
  if ((sum(is.na(data_train[ , i]))/160) >= 50) {
    predictors <- predictors[!(predictors == names(data_train)[i])]
  }
}

data_train_clean <- select(data_train, one_of(predictors, "classe"))
```

### 2.3 Train and select a prediction model
First of all a randomn forest model is trained. This model type is known for good performance on classification problems. In order to evaluate the trained model 3-fold cross validation is used to estimate the out of sample error.

```{r, cache = TRUE}
set.seed(741)
mod_fit <- train(classe ~ .,
                 data_train_clean,
                 method = "rf",
                 ntree = 50,
                 trControl = trainControl(method = "cv", 
                                          number = 3))

  confusionMatrix(mod_fit)
```

The averaged accuracy of the cross validated confusion matrix is already sufficient. Therefore, the trained model is suitable.

### 2.4 Out of sample error
In order to estimate the out of sample error the model is used to predict the test sample and compared to the correct results. 

```{r}
predict_test <- predict(mod_fit, data_test)
confusionMatrix(predict_test, data_test$classe)[2:3]
```

The out of sample error of the trained model on the test set is quite low (1-Accuracy). The model seems to work very well.

## Conclusion
The trained random forest model accurately classifies the way the participants performed the excercise with the provided data.

## Annex

### Environment
```{r}
sessionInfo()
```

